---
title: "extractr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{extractr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
#| label: not_include

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
# devtools::install_local(force = T)
# if (interactive())
#   devtools::load_all()
unlink(here::here("data_tmp"), recursive = TRUE)
```

Load packages and set up the environment.

```{r}
#| label: load_extractr
#| eval: true

library(extractr)
```

```{r}
#| label: load_pkgs

suppressPackageStartupMessages({
  library(dplyr)
  library(here)
  library(mapview)
  library(readr)
  library(sf)
  library(terra)
})
options(readr.show_col_types = F)
```

## Get area of interest (AoI)

Load area of interest (AoI) from this [sanctuaries.geojson](https://github.com/noaa-onms/onmsR/blob/master/data-raw/sanctuaries.geojson) for the Florida Keys National Marine Sanctuary (FKNMS).

```{r}
#| label: aoi

aoi_geo <- "https://raw.githubusercontent.com/noaa-onms/onmsR/master/data-raw/sanctuaries.geojson"

aoi <- read_sf(aoi_geo) |> 
  filter(nms == "FKNMS")
(bb <- st_bbox(aoi))

mapView(aoi)
```

## Dataset: sea surface temperature

### Get info

ERDDAP dataset:

- [ERDDAP - Sea Surface Temperature, NOAA Coral Reef Watch Daily Global 5km Satellite SST (CoralTemp), 1985-present, Daily - Data Access Form](https://coastwatch.noaa.gov/erddap/griddap/noaacrwsstDaily.html)

```{r}
#| label: sst_ed_info

ed_url  <- "https://coastwatch.noaa.gov/erddap/griddap/noaacrwsstDaily.html"
(ed <- ed_info(ed_url))
```

### Get dimensions

```{r}
#| label: sst_ed_dims

var <- "analysed_sst"

dims <- ed_dims(ed)
names(dims)

# let's limit to the most recent 10 times
(times <- tail(dims[["time"]], 10))
```

### Extract dataset from AoI

```{r}
#| label: sst_ed_extract

d_csv <- here("data_tmp/sst_timeseries.csv")
r_tif <- here("data_tmp/sst_raster.tif")

(d <- ed_extract(
  ed        = ed,
  var       = var,
  sf_zones  = aoi,
  fld_zones = "nms",
  zonal_fun = "mean",
  zonal_csv = d_csv,
  rast_tif  = r_tif,
  time_min  = times[1],
  verbose   = TRUE))
```

### Plot time series

```{r}
#| label: sst_plot_ts

d <- read_csv(d_csv)
head(d)

plot_ts(d, label = "Surface Temperature (ºC)")
```

### Map raster

```{r}
#| label: sst_map_rast

r <- rast(r_tif)
names(r)

lyr <- names(r)[1]
plet(r[lyr], tiles = "Esri.OceanBasemap")
```


## Dataset: sea surface salinity

### Get info

ERDDAP dataset:

- [ERDDAP - Sea Surface Salinity, Miras SMOS, Near Real-Time, Global 0.25°, 2010-present, 3 Day Composite - Data Access Form](https://coastwatch.noaa.gov/erddap/griddap/noaacwSMOSsss3day.html)

```{r}
#| label: sss_ed_info

ed_url  <- "https://coastwatch.noaa.gov/erddap/griddap/noaacwSMOSsss3day.html"
ed <- ed_info(ed_url)
ed
```

### Get dimensions

```{r}
#| label: sss_ed_dims

var <- "sss"

dims <- ed_dims(ed)
names(dims)

# let's limit to the most recent 10 times
(times <- tail(dims[["time"]], 10))
```

### Extract dataset from AoI

```{r}
#| label: sss_ed_extract

d_csv <- here("data_tmp/sss_timeseries.csv")
r_tif <- here("data_tmp/sss_raster.tif")

(d <- ed_extract(
  ed        = ed,
  var       = var,
  sf_zones  = aoi,
  fld_zones = "nms",
  zonal_fun = "mean",
  zonal_csv = d_csv,
  rast_tif  = r_tif,
  time_min  = times[1],
  verbose   = TRUE))
```

### Plot time series

```{r}
#| label: sss_plot_ts

d <- read_csv(d_csv)
head(d)

plot_ts(d, label = "Surface Salinity (PSU)")
```

### Map raster

```{r}
#| label: sss_map_rast

r <- rast(r_tif)
names(r)

lyr <- names(r)[1]
plet(r[lyr], tiles = "Esri.OceanBasemap")
```
