---
title: "extractr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{extractr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
# devtools::install_local(force = T)
# devtools::load_all()
```

```{r setup}
library(extractr)
```

## Get ERDDAP dataset

ERDDAP dataset info and date ranges for two available datasets: global_monthly and global_8day.

```{r}
ed_url  <- "https://apdrc.soest.hawaii.edu/erddap/griddap/hawaii_soest_d749_a206_cd3a.html"
ed <- get_ed_info(ed_url)
ed
```

```
<ERDDAP info> hawaii_soest_d749_a206_cd3a 
 Base URL: https://apdrc.soest.hawaii.edu/erddap 
 Dataset Type: griddap 
 Dimensions (range):  
     time: (1992-01-02T00:00:00Z, 2023-04-28T00:00:00Z) 
     LEV: (5.0, 5906.25) 
     latitude: (-89.875, 89.875) 
     longitude: (0.125, 359.875) 
 Variables:  
     salt: 
         Units: PSU 
```

Additional ERDDAP datasets of interest:
- https://upwell.pfeg.noaa.gov/erddap/griddap/jplMURSST41mday.html
- https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41mday.html

## Get area of interest

Load area of interest (AoI) geojson files.

```{r}
library(dplyr)
library(mapview)
library(sf)

aoi_geo <- "https://raw.githubusercontent.com/noaa-onms/onmsR/master/data-raw/sanctuaries.geojson"

aoi <- sf::read_sf(aoi_geo) |> 
  dplyr::filter(nms == "FKNMS")

mapview::mapView(aoi)
```

## Extract

Extract a time-series for a given dataset and AoI

```{r extract-old}
#| eval: false

dir_tif <- here::here("data_tmp")  # directory for storing .tif files
grds <- get_ed_grds(
  ed_info, 
  ed_var="sst",  # variable name from the ERDDAP 
  ply=fk_sf,  # polygon of the RoI
  dir_tif=dir_tif, 
  date_beg = "2021-10-01"  # start of time-series
  # date_end defaults to the current date
)
```

## Get datetimes

- Fix dimensions

```{r}
# arguments in
var    <- "salt"
var_z  <- "LEV"
n_max  <- 10000

rerddap:::dimvars(ed)
# "time"      "LEV"       "latitude"  "longitude"
ts <- get_ed_dim(ed, "time")      # 3,814
zs <- get_ed_dim(ed, "LEV")       #    50
xs <- get_ed_dim(ed, "longitude") # 1,440
ys <- get_ed_dim(ed, "latitude")  #   720

# let's limit to the most recent 10 times
n_t <- length(ts) 
ts[(n_t - 10):n_t]
```


## Summarize the Extracted Time Series

```{r}
ts_csv <- here::here("data_tmp/ts.csv")
ts <- grds_to_ts(
  grds, fxns = c("mean", "sd"), ts_csv, verbose = T)
head(ts)
```

## Plot the Extracted Time Series
```{r}
plot_ts(ts_csv, "mean", "sd")
```
